---
title: "P8105_HW3_jao2195"
author: "Jennifer O."
date: "2023-10-10"
output: github_document
---

################################################################################
### Problem 0 
This “problem” focuses on structure of your submission, especially the use git and GitHub for reproducibility, R Projects to organize your work, R Markdown to write reproducible reports, relative paths to load data from local files, and reasonable naming structures for your files.

To that end:

-create a public GitHub repo + local R Project; we suggest naming this repo / directory p8105_hw3_YOURUNI (e.g. -p8105_hw3_ajg2202 for Jeff), but that’s not required
-create a single .Rmd file named p8105_hw3_YOURUNI.Rmd that renders to github_document
-create a subdirectory to store any local data files, and use relative paths to access these data files
-submit a link to your repo via Courseworks
-Your solutions to Problems 1, 2, and 3 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems.

For this Problem, we will assess adherence to the instructions above regarding repo structure, git commit history, and whether we are able to knit your .Rmd to ensure that your work is reproducible. Adherence to appropriate styling and clarity of code will be assessed in Problems 1+ using the style rubric.

This homework includes figures; the readability of your embedded plots (e.g. font sizes, axis labels, titles) will be assessed in Problems 1+.

################################################################################
### Problem 1
This problem uses the Instacart data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets using:

```{r,message=FALSE}
# Loading My Libraries
library(readxl)
library(ggplot2)
library(tidyverse)
library(readr)
library(lubridate)
library(magrittr)
library(dplyr)
library(ggridges)
library(patchwork)
library(p8105.datasets)

```

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

```{r}
#Loading Instacart Data 
library(p8105.datasets)
data("instacart")

```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations. Then, do or answer the following (commenting on the results of each):

How many aisles are there, and which aisles are the most items ordered from?
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
#Renaming main dataset
insta = instacart

#class(insta)
#Number of Aisles
# View(insta)
num_aisles = nrow(unique(insta$aisle))
num_aisles
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

################################################################################
### Problem 2

#### Problem [2.1] - This problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package using

```{r}
# Loading brfss data directly; not storing to local directory
library(p8105.datasets)
data("brfss_smart2010")
```

#### Problem [2.2] - First, do some data cleaning: 1) Format the data to use appropriate variable names;focus on the “Overall Health” topic include only responses from “Excellent” to “Poor”. 2) Organize responses as a factor taking levels ordered from “Poor” to “Excellent”.

```{r}
#Formatting data using appropriate variable names; Topic "Overall Health".

brfss = brfss_smart2010 #Renaming Dataset to shorter name 


brfss = brfss %>% #Overwritting dataset
        janitor::clean_names() %>%  #Cleaning Dataset / checking appropriate variable names
        filter(topic == "Overall Health") %>%  #Filtering Dataset down to Overall Health Topic
        filter(response == "Excellent" | response == "Very good"| response == "Good" | response =="Fair" | response =="Poor") %>%         #Selecting Responses that were on the scale of Excellent down to Poor
        mutate(response=fct_relevel(response,c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% 
        #Organizing the factor level, ordered from Poor to Excellent, where Poor Factor Level == 1 
        #and Excellent Factor Level == 5.
        rename(state = locationabbr) %>% #Renaming to more appropriate variable name
        rename(state_county = locationdesc) #Renaming to more appropriate variable name

```

#### Problem [2.3] Using this dataset, do or answer the following (commenting on the results of each): In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
#Determining States Observed at 7 or more locations in 2002.
brfss_states_2002 = brfss %>% #Piping 
  group_by(state, year) %>% #Grouping by state, year.
  summarize(n_locations = n()) %>% #Summarizing 
  filter(n_locations >= 7 & year== 2002) #Filtering down to >7 and year == 2002 criteria to answer question.

#Determining States Observed at 7 or more locations in 2010.
brfss_states_2010 = brfss %>% #Piping 
  group_by(state, year) %>% #Grouping by state, year.
  summarize(n_locations = n()) %>% #Summarizing 
  filter(n_locations >= 7 & year== 2010) #Filtering down to >7 and year == 2010 criteria to answer question.

#Listing States that meet the requested criteria:
states2002 = select(brfss_states_2002,state)
states2010 = select(brfss_states_2010,state)

# There are `r nrow(states2002)` states that were observed at 7 or more locations in 2002. 
# The `r nrow(states2002)` states in 2002 were: `r states2002`.
# There are `r nrow(states2010)` states that were observed at 7 or more locations in 2010. 
# The `r nrow(states2010)` states in 2010 were: `r states2002`.
# There are `r nrow(states2010) - nrow(states2002)` more states that meet this criteria in 2010, than in 2002. 
```
#### Problem [2.3] Comments on the Results of the Data:




#### Problem [2.4] Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent_only <- brfss %>% 
  filter(response=="Excellent") %>% 
  group_by(year, state) %>% 
  summarize(average_data_value = mean(data_value, na.rm = TRUE))


excellent_spaghetti_plot = ggplot(excellent_only, 
                           aes(x=year, y=average_data_value, color=state, group=state))+
                           geom_point()+
                           geom_line()+
                           labs(title = "Plot:The Average Value By State from 2002 to 2010",
                           x = "Years",
                           y = "Average data value",
                           color = "States",
                           )+
                           theme(legend.position = "left")

excellent_spaghetti_plot
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}

brfss_panel <- brfss %>% 
  filter(state=="NY" & year %in% c(2006,2010)) 


# myplot = ggplot(brfss_panel, aes(x = data_value, fill = response)) +
#   geom_density(alpha = .) + 
#   facet_grid(~ year) + 
#   viridis::scale_fill_viridis(discrete = TRUE)+
#   labs(title = "Poor” to “Excellent” Distribution of data_value",
#        x="Data Value",
#        y="Density")
# 
# myplot
```

################################################################################
### Problem 3

Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices can measure MIMS in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses accelerometer data collected on 250 participants in the NHANES study. The participants’ demographic data can be downloaded here, and their accelerometer data can be downloaded here. Variables *MIMS are the MIMS values for each minute of a 24-hour day starting at midnight.

### Problem [3.1] - Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).


```{r}
#Importing Nhanes Dataset Covar
nhanes_covar =  read_csv("nhanes_covar.csv",
    skip = 4, na = "NA") %>%  #Loading data
    janitor::clean_names() %>%  #Tidying Data
    drop_na() %>% #Excluding participants with missing demographic data (with NAs) in their cells
    filter(age >= 21) %>% #Kepping participants that are greater than or equal to 21 yrs
    mutate(education = factor(education, levels = c(1, 2, 3), labels = c("Less than high school", "High school equivalent", "More than high school"))) %>% #Creating Factor levels for education
 mutate(sex = ifelse(sex == 1, "Male", "Female")) # Changing sex from numeric to character 

# Importing Nhanes Dataset Accel
nhanes_accel =  read_csv("nhanes_accel.csv") %>% #Imporing Accelerometer data 
                janitor::clean_names() # Cleaning Names
  
# Merge NHANES COVAR & NHANES ACCEL
 merged_covar_accel= left_join(nhanes_covar, nhanes_accel, by= "seqn")

```


### Problem [3.2]  Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}
#Reader Friendly Table 
mytable = merged_covar_accel %>% 
  select(education,sex) %>% 
  group_by(education, sex) %>% 
  summarize(n_obs = n()) %>%  
  knitr::kable(digits = 3)

mytable

#Visualization of the Age distributions 
age_dist<-merged_covar_accel %>% 
ggplot(aes(x=age , y=education, fill=sex), )+
  geom_density_ridges(scale=.85, alpha=0.4)+
  labs(title = "Age Distribution by Education and Sex",
    x = "Age",
    fill = "Sex",)
  
  age_dist
  

```
### Problem [3.2] Description of Friendly Table 

Overall, we started with 250 participants. After exclusion criteria was applied, there were 228 partiicpant that were included for data anlysis. In the first education group of 1) LESS THAN A HIGH SCHOOL EDUCATION there is about and EQUAL number of males (28) and females (27). In the second education group 2)  HIGH SCHOOL EQUIVALENT there is approximately 1.5 more males (35) than females (23). In the last education group, 3) MORE THAN HIGH SCHOOL, there are roughly and EQUAL amount of males (56) to females (59). The larges discrepancy between males and females is betweens was found in the 2) HIGH SCHOOL EQUIVALENT group. 


### Problem [3.2] Description of Visualization of the Age distributions 

hhhhhhhhh 


### Problem [3.3]
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}

merged_covar_agg <- merged_covar_accel %>% #Taking merged dataset
                    mutate(total_activityee = rowSums(select(merged_covar_accel,starts_with("min")))) %>% #Summing across minutes
                    relocate(seqn, sex, age, bmi, education, total_activityee, everything()) #Relocating coulums to the front of dataset


totact_plot = ggplot(merged_covar_agg , aes(x = age, y = total_activityee, color = sex, group=sex)) +
  geom_point(alpha =0.6) +
  geom_smooth(se = FALSE) +
  facet_grid(~ education) +
  theme(legend.position = "left")+
  labs(title = "Total Activity Across Age Groups",
       x="Age",
       y="Total Activity")

totact_plot 


#%  mutate(merged_covar_accel , total_activ = rowSums(select(.,starts_with("min")))) %>% 

```

### Problem [3.4]
Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}

merged_final <- merged_covar_agg %>% 
mutate(merged_covar_agg, hr_activity= total_activityee/60) %>% 
relocate(seqn, sex, age, bmi, education, total_activityee, hr_activity, everything())


my24hrplot =ggplot(merged_final, aes(x = hr_activity, fill = sex)) +
  geom_density(alpha = .5) + 
  facet_grid(~education) + 
  viridis::scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "bottom")

my24hrplot

```


